Paper 2: survey analysis

### 2024-09-05 Rmd/Paper2.Rmd
Fixed and tidied the presentation of the results.

---
title:        "Paper 2: survey analysis"
author:       "Niklas Karlsen"
date:         "2024-09-04"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

@@

@@

For the survey, the complete TRELIS-Q dataset was used.

@@

```{r load_data_cfa}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
```

@@

OBS! Anonymiser data i datafil på studiested før det legges ut åpent på nett

@@

```{r fix_missing, eval=T}
# remove NMAR + Institute9 (19)
data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),]

# jeg skjuler Annet, siden det bare er en person; alternativt NA?
data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
#trinn_r[is.na(trinn_r)] <- 1

@@

The purpose of the multiple regression analysis is to explain the variation in the data. I therefore use stepwise regression and add the variables step-by-step to examine changes in R2 (explained variance).

lm_robust is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)

@@

@@

@@

@@

TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet. (Jeong and Jung, 2016)

@@

@@

@@

## Reanalysis with interaction effects omitted

@@

@@

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TK_covariates_l1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(model_TK_covariates_l1)
```

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center = T, scale = T) ~ factor(Gender) * (factor(Trinn) + factor(semester)),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

@@

```{r TKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TK_covariates_l2 <- lm_robust(
  scale(TK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TK_covariates_l2)
```

```{r TKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TK) ~
    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

@@

@@

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_PCK_covariates_l1 <- lm_robust(PCK ~ factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l1)
```

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center = T, scale = T) ~ 
                  factor(Trinn) * (factor(Gender) + factor(semester)) +
                  factor(semester) * factor(Gender),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

@@

```{r PCKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_PCK_covariates_l2 <- lm_robust(
  scale(PCK) ~ 
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_covariates_l2)
```

```{r PCKlevel2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ 
                  factor(Gender) + factor(Trinn) + factor(semester) +        # level 1
                  factor(studiested) * (factor(semester) + factor(Gender)),  # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

@@

@@

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TPACK_covariates_l1 <- lm_robust(
  TPACK ~
    factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l1)
```

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(
  scale(TPACK, center = T, scale = T) ~
    factor(Gender) * factor(semester) + factor(Trinn),
  clusters = factor(studiested), data = data_clean
)
summary(m2)
```

@@

```{r level2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TPACK_covariates_l2 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_covariates_l2)
```

```{r level2_interaction}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) +       # level 1
    factor(studiested) * (factor(semester) + factor(Gender)), # level 2
  clusters = factor(studiested), data = data_clean
)
summary(m4)
```

@@

```{r PCK}
# PCK
model_TPACK_PCK <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested) +                                # level 2
    scale(PCK),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_PCK)

@@

```{r TK}
# TK
model_TPACK_TK <- lm_robust(
  scale(TPACK) ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested) +                                # level 2
    scale(PCK) + scale(TK),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_TK)

@@

@@

@@

@@

@@

@@

@@

@@

The results are discussed in @karlsen_assessing_2024.

>>>>

---
title:        "Paper 2: survey analysis"
author:       "Niklas Karlsen"
date:         "2024-09-05"
bibliography: "../data/meta_data/references.bib"
csl:          "../scripts/university-of-gothenburg-apa-7th-edition-swedish-legislations.csl" # (source: https://bookdown.org/yihui/rmarkdown-cookbook/bibliography.html)
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true

@@

Since the data is structured at two levels (i.e., PSTs and TEIs), the data is not independent as is assumed when doing OLS regression. Since only 5 TEIs are in the dataset, which is below the recommended miminum of 10, OLS with cluster corrected standard errors may be a feasible option to MLM.

The ICC is checked to see how much of the variance is accounted for by the level 2 units, to assess whether it is necessary to account for the level 2-dependence. ICC should be above 5 % for MLM to be considered.

The OLS regression is hierarchical. The covariates are introduced block-wise. Gender and target grade is related to the individual PST, while year of study can reflect both changes at level 1 and level 2.

For the OLS the TEIs are introduced as dummy variables, as are the other covariates with more than two levels.

For the multi-level modeling (MLM) the PSTs are considered level 1 units and the TEIs are level 2. The covariates gender and target grade are related to the level 1 units, while year of study can vary at both levels.

In addition, TK and PCK are considered independent variables, while TPACK is the dependent variable.

Both the OLS and the MLM are done by introducing the variables sequentially. For the OLS done in paper 2, the level 1 variables (i.e., gender, trinn and semester) are introduced first, then the level 2 variables (i.e., TEIs), and finally, PCK and TK. For the MLM, a baseline model (null-model) is created first, where the TEIs are set as random (i.e., the intercept is allowed to vary among TEIs). Then the level 1 variables are introduced to create the level 1-model. Then PCK and TK are added to the model. Since the order of introducing TEI was altered between the OLS in paper 2 and the MLM, a reanalysis of the OLS was done with the TEIs introduced before the covariates. This could affect how much of the variance was explained by the covariates.

The explained variance (i.e., R2) is calculated for both methods, and can be used to compare the results of the two methods.

The following models are tested:

Model | Paper 2 OLS | Reanalysis OLS | New analysis MLM 
------------- | ------------- | ------------- | ------------- 
Options | 3xAnnet = NA | N/A | N/A
Nullmodell | N/A | TPACK ~ TEIs | TPACK ~ 1
Level 1 | TPACK ~ covariates + TEIs | TPACK ~ TEIs + covariates | TPACK ~ covariates
Level 2 | N/A | N/A | + year of study | TEI
PCK | TPACK ~ covariates + TEIs + PCK | TPACK ~ TEIs + covariates + PCK | TPACK ~ covariates + PCK_sentrert
TK | TPACK ~ covariates + TEIs + PCK + TK | TPACK ~ TEIs + covariates + PCK + TK | TPACK ~ covariates + PCK_s + TK_sentrert

The analysis in paper 2 changed the answers of three respondents from "Other" to NA. This made it impossible to compare the fit for the MLMs since the number of respondents were dissimilar for the models compared when including these covariates. For the reanalyses I therefore kept these values unchanged.

In the MLM the TEIs are added as random variables (i.e., the intercepts are allowed to differ between TEIs), while the TEIs are added as dummy variables in the OLS. This should give similar results.

In the MLM the covariates are added as fixed variables (i.e., the slopes are the same for all PSTs independent of TEI).

The order of the models was also changed in the reanalysis, as the baseline model in the MLM corresponds to having the TEIs as dummy variables. In paper 2 this model was tested after having controlled for the covariates. In the reanalysis this order was switched to correspond to the MLM.

Only the MLM has the option to test for level 2 covariates. I allowed slope of year of study vary between TEIs. The level2 model is therefore not tested. It did not converge.

For the MLM the independendent variables were centred using the grand mean, following @christophersen_introduksjon_2018.

The different methods can be compared by comparing R-squared, and diff R-squared.

@@

For the survey, the complete TRELIS-Q dataset was used. Here I load the raw data and prepare it for analysis.

@@

The environment is reset and the data reloaded.

```{r load_data}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
```

@@

@@

```{r fix_missing, eval=T}
# remove NMAR + Institute9 (19)
data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
```

To reproduce paper 2 values, I set the answers "Other" to NA.

```{r}
# jeg skjuler Annet, siden det bare er en person; alternativt NA?
data_clean$Trinn[data_clean$Trinn == "Annet"] <- NA # "Trinn 5-10"
#trinn_r[is.na(trinn_r)] <- 1

@@

The purpose of the multiple regression analysis is to explain the variation in the data. I therefore use hierarchical regression [@howitt_introduction_2014] and add the variables step-by-step to examine changes in R2 (explained variance).

@@

lm_robust [@blair_estimatr_2022] is used to calculate cluster corrected standard errors. (or robust SE? Actually, CR2 SE = cluster robust standard errors)

@@

The residuals need to be normally distributed for the statistical tests. For dichotomous variables, the distribution should be less than 80/20 skewed.

@@

The data should be linearly related.

@@

TK korrelererer veldig sterkt (0.84) med TPACK, noe som indikerer kollinearitet (Jeong and Jung, 2016). This issue is discussed in @karlsen_assessing_2024 and the synopsis.

@@

Here I reproduce the results reported in paper 2. Interaction effects were also tested to see if they needed to be included in the analysis.

@@

Here I reproduce the tables presented in paper 2. The significance values can be found in the f-tests done above.

@@

## Reanalysis of OLS

A reanalysis of the OLS is done here, due to some methodical reconsiderations after a closer look at the MLM. The dataset is slightly closer to the rawdata (3 NAs were reset to other, as in the original dataset). The order of the variables are also slightly different (TEIs are introduced before covariates). Finally, interaction effects are not included, as they were in paper 2. The statistical tests suggested that the interaction effects were non-significant. Otherwise, the analysis is similar to paper 2.

### Reload data

```{r}
# ref: https://www.statology.org/clear-environment-in-r/ [5.9.2024]
#clear all data frames from environment
rm(list=ls(all=TRUE)[sapply(mget(ls(all=TRUE)), class) == "data.frame"])
```

Then I reload the dataset from file.

```{r load_data_reanalysis}
# Rdata-file contains reduced / prepared datasets
load(file="../data/processed_data/tpack-TRELIS-Q_complete_2023-paper2.Rdata")
```

Then I recenter the data.

And add some missing information.

NMAR are removed.

```{r fix_missing_2, eval=T}
# remove NMAR + Institute9 (19)
data_clean <- data_TRELIS_Q[-c(19,34,44,45,46,95,102),] 
```

@@

#### baseline model (nullmodell)

```{r TK_baseline}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TK_baseline <- lm_robust(TK ~ factor(studiested), 
                clusters = factor(studiested), data=data_clean)
summary(model_TK_baseline)
```

@@

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TK_level1 <- lm_robust(TK ~ factor(studiested) + factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(model_TK_level1)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- model_TK_level1$adj.r.squared / model_TK_baseline$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

The F-statistic indicate that the two models are not significantly different. This means that level1 effects do not contribute significantly to the variance.

@@

#### Baseline

```{r PCKlevel2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_PCK_baseline <- lm_robust(
  PCK ~ 
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_baseline)
```

@@

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_PCK_level1 <- lm_robust(PCK ~ factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_PCK_level1)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- model_PCK_level1$adj.r.squared / model_PCK_baseline$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There were significant differences when including covariates.

@@

#### Baseline

```{r level2_main}
# studiested; Institute1 velges som referansekategori siden den har lavest TPACK-verdi av alle
model_TPACK_baseline <- lm_robust(
  TPACK ~
    factor(studiested),                                 # level 2
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_baseline)
```

@@

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
model_TPACK_level1 <- lm_robust(
  TPACK ~
    factor(studiested) + factor(Gender) + factor(Trinn) + factor(semester),
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_level1)
```

@@

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- model_TPACK_level1$adj.r.squared / model_TPACK_baseline$adj.r.squared
df1 <- length(data_clean$TPACK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

There is no significant difference when including covariates.

@@

```{r PCK}
# PCK
model_TPACK_PCK <- lm_robust(
  TPACK ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested) +                                # level 2
    PCK,
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_PCK)

# main
fstat <- model_TPACK_PCK$adj.r.squared / model_TPACK_level1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

@@

```{r TK}
# TK
model_TPACK_TK <- lm_robust(
  TPACK ~
    factor(Gender) + factor(Trinn) + factor(semester) + # level 1
    factor(studiested) +                                # level 2
    PCK + TK,
  clusters = factor(studiested), data = data_clean
)
summary(model_TPACK_TK)

@@

mtkcl1rs  <- model_TK_baseline$r.squared
mtkcl1ars <- model_TK_baseline$adj.r.squared

mtkcl2rs  <- model_TK_level1$r.squared
mtkcl2ars <- model_TK_level1$adj.r.squared

@@

rownames(TK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")

@@

mpcl1rs   <- model_PCK_baseline$r.squared
mpcl1ars  <- model_PCK_baseline$adj.r.squared

mpcl2rs   <- model_PCK_level1$r.squared
mpcl2ars  <- model_PCK_level1$adj.r.squared

@@

rownames(PCK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level")

@@

mtcl1rs  <- model_TPACK_baseline$r.squared
mtcl1ars <- model_TPACK_baseline$adj.r.squared

mtcl2rs  <- model_TPACK_level1$r.squared
mtcl2ars <- model_TPACK_level1$adj.r.squared

@@

rownames(TPACK_R2) <- c("Teacher education institute", "Gender, year of study, target grade level", "PCK", "TK")

@@

When the data is nested, multi-level modeling can be used. However, when the number of level 2-units is below 10, as it is in our dataset (we have only 5 TEIs), OLS with cluster corrected standard errors can be used. I, however, do a MLM here to compare with the OLS results. MLM is only done for TPACK as dependent variable to illustrate the method.

The r package NLME [@pinheiro_nlme_2023] is used for the analysis.

### Baseline model

First, the baseline model is calculated as a reference and to find the ICC.

```{r}
# baseline model / nullmodell

# REML is used for parameter estimation
model_MLM_baseline <- lme(TPACK~1, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)

# ML is used for model fit
model_MLM_baseline_fit <- lme(TPACK~1, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)

summary(model_MLM_baseline)
summary(model_MLM_baseline_fit)
```

```{r}
# ICC
nu_00 <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
nu_e  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)
ICC   <- nu_00 / (nu_00 + nu_e)
ICC
```

30 % av variansen i TPACK ligger på nivå 2. Dette er større enn den anbefalte grensen på 5 %. Vurder om OLS-regresjon med klyngekorrigert standardfeil egner seg bedre enn flernivåanalyse [@christophersen_introduksjon_2018] (p. 109).

```{r}
# effektiv utvalgsstørrelse
n_e <- 147 / (1 + ((147 / 5) - 1) * ICC)
n_e
```

Utvalgsstørrelsen korrigert for avhengighet er lik 15.

### Level 1 model (random intercept)

The level 1 model extends the baseline model by including the covariates.

#### Level 1 variables

```{r}
# REML is used for parameter estimation
model_MLM_level1 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)

# ML is used for model fit
model_MLM_level1_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)

summary(model_MLM_level1)
summary(model_MLM_level1_fit)
```

Er level 1-modellen bedre enn nullmodellen?

```{r eval=T}
anova(model_MLM_baseline_fit, model_MLM_level1_fit)
```

#### Level 2 variables

Since Level 2 variables are unique to MLM, and the solution does not converge, level 2 analysis is not done here.

La kovariatene variere mellom studiesteder

```{r eval=F}
# REML is used for parameter estimation
model_level2 <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="REML", na.action="na.omit", random= ~ semester|studiested)

# ML is used for model fit
model_level2_fit <- lme(TPACK~Gender + semester + Trinn, data=data_clean, method="ML", na.action="na.omit", random= ~ semester|studiested)

summary(model_level2)
summary(model_level2_fit)
```

Er level 2-modellen bedre enn level 1?

```{r eval=F}
anova(model_level2_fit, model_MLM_level1_fit)
```

### PCK

```{r}
# grand mean centering: scale(TK, center = T, scale = F)

data_clean$PCK_cent <- scale(data_clean$PCK, center = T, scale = F)
data_clean$TK_cent  <- scale(data_clean$TK, center = T, scale = F)

# REML is used for parameter estimation
model_MLM_PCK <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)

# ML is used for model fit
model_MLM_PCK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)

summary(model_MLM_PCK)
summary(model_MLM_PCK_fit)
```

```{r eval=T}
anova(model_MLM_PCK_fit, model_MLM_level1_fit)
```

### TK

```{r}
# REML is used for parameter estimation
model_MLM_TK <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="REML", na.action="na.omit", random= ~1|studiested)

# ML is used for model fit
model_MLM_TK_fit <- lme(TPACK~Gender + semester + Trinn + PCK_cent + TK_cent, data=data_clean, method="ML", na.action="na.omit", random= ~1|studiested)

summary(model_MLM_TK)
summary(model_MLM_TK_fit)
```

```{r eval=T}
anova(model_MLM_PCK_fit, model_MLM_TK_fit)
```

## Comparison of OLS and MLM

### Comparison of variance

I follow @christophersen_introduksjon_2018 (p. 115) and use the variance estimates as an indicator of R2.

```{r}
# MLM
# variance baseline
nu_00_baseline <- var(ranef(model_MLM_baseline)) # intercept (i.e., TEIs)
nu_e_baseline  <- var(resid(model_MLM_baseline)) # residuals (i.e., PSTs)

# variance level 1
nu_00_level1 <- var(ranef(model_MLM_level1)) # intercept (i.e., TEIs)
nu_e_level1  <- var(resid(model_MLM_level1)) # residuals (i.e., PSTs)

# variance PCK
nu_00_PCK <- var(ranef(model_MLM_PCK)) # intercept (i.e., TEIs)
nu_e_PCK  <- var(resid(model_MLM_PCK)) # residuals (i.e., PSTs)

# variance TK
nu_00_TK <- var(ranef(model_MLM_TK)) # intercept (i.e., TEIs)
nu_e_TK  <- var(resid(model_MLM_TK)) # residuals (i.e., PSTs)

# OLS
mtcl1ars <- model_TPACK_baseline$adj.r.squared
mtcl2ars <- model_TPACK_level1$adj.r.squared
mtpars   <- model_TPACK_PCK$adj.r.squared
mttars   <- model_TPACK_TK$adj.r.squared

# table of difference in variance

table_variance <- data.frame(
  c(nu_e_baseline, (nu_e_baseline - nu_e_level1) / nu_e_baseline, (nu_e_level1 - nu_e_PCK) / nu_e_level1, (nu_e_PCK - nu_e_TK) / nu_e_PCK),
  c(nu_00_baseline, (nu_00_baseline - nu_00_level1) / nu_00_baseline, (nu_00_baseline - nu_00_PCK) / nu_00_baseline, (nu_00_baseline - nu_00_TK) / nu_00_baseline),
  c(mtcl1ars, mtcl2ars - mtcl1ars, mtpars - mtcl2ars, mttars - mtpars)
  )

rownames(table_variance) <- c("TEI", "covariates", "PCK", "TK")
colnames(table_variance) <- c("MLM Level 1 % var", "MLM Level 2 % var", "OLS reanalysis diff R2")

pander(table_variance)
```

The variance differ somewhat from the standard deviation (when squared) reported in the summary of the MLMs. Why? Is the OLS a kind of average of the MLM, which is separated into level 1 and level 2?

@@

I was able to replicate the results discussed in @karlsen_assessing_2024.

The reanalysis using OLS gave similar results as for paper 2. Only TEIs explained variance in TK, not the covariates. The covariates did however significantly explain the variance in PCK, but the TEIs did not (or to a small degree, only 3%). The variance in TPACK was significantly explained by TEI, PCK and TK, not the covariates.

The results in paper 2 indicated that the covariates contributed to explain the variance in TPACK. The reanalysis done here, suggest that this may be due to the order in which the variables were added to the model being tested. If I follow the order used in the MLM, the reanalysis suggests that the covariates do not contribute to TPACK.

The MLM gave supporting results. Including the covariates did not improve the model, whereas including PCK and TK did. The variance at level 2 (between TEIs) was due to TK, not PCK or covariates, that is, there were significant differences in TK between TEIs. At level 1 (between PSTs) 6% of the variance were due to the covariates, 17 % due to PCK and 55 % due to TK.

The OLS reanalysis suggested that 30 % of the variance in TPACK was due to TEIs, 11 % due to PCK and 32 % due to TPACK. The covariates did not contribute to the variance in TPACK. Even though the percentages differ between the OLS and the MLM, they show the same trends. The different percentages may be due to how R2 (or difference in variance) is calculated differently in OLS and MLM.
